RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-06 paper/RDF validation requirements/RDF validation requirements.tex
solved some issues

\begin{itemize}
	\item R-33-NEGATIVE-OBJECT-CONSTRAINTS
	\item R-200-NEGATIVE-LITERAL-CONSTRAINTS
\end{itemize}

\begin{itemize}
	\item not allowed objects
	\item not allowed literals
	\item definition: A matching triple has any literal / object except those explicitly excluded
\end{itemize}

@@

\begin{itemize}
	\item Siths do not have blue, green, or white laser swords
\end{itemize}

\begin{ex}
# OWL2
# ----
:laserSwordColor rdfs:range :negativeLaserSwordColors . 
:NegativeLaserSwordColors
    a rdfs:Datatype .
    owl:complementOf :laserSwordColors .
:laserSwordColors
    a rdfs:Datatype .
    owl:oneOf ( "blue" "green" "white" ) .
\end{ex}

\begin{ex}
# ShEx
# ----
:Sith {
    ! :laserSwordColor ('blue' 'green' 'white') }
\end{ex}

\begin{ex}
# Sith individuals
# ----------------
:DarthSidious
    :laserSwordColor 'red' .
\end{ex}

@@

\begin{itemize}
	\item R-32-MEMBERSHIP-OF-RDF-OBJECTS-IN-CONTROLLED-VOCABULARIES
	\item R-39-MEMBERSHIP-OF-RDF-LITERALS-IN-CONTROLLED-VOCABULARIES
\end{itemize}

\begin{itemize}
	\item resources can only be members of listed controlled vocabularies
\end{itemize}

@@

\begin{ex}
# DSCL (DSP)
# ----------
:bookDescriptionTemplate 
    a dsp:DescriptionTemplate ;
    dsp:resourceClass swrc:Book ; 
    dsp:statementTemplate [
        a dsp:NonLiteralStatementTemplate ;
        dsp:property dcterms:subject ; 
        dsp:nonLiteralConstraint [ 
            a dsp:NonLiteralConstraint ;
            dsp:valueClass skos:Concept ; 
            dsp:vocabularyEncodingScheme :BookSubjects, :BookTopics, :BookCategories ] ] .
\end{ex}

\begin{itemize}
	\item \ms{skos:Concept} resources can only be members of the listed controlled vocabularies
\end{itemize}
%\er{This sort of DL-syntax (with semicolon) must be rechecked. Note that in DL, it does not have to correspond directly to names of things in machine-readable syntax. It is fine to say BookSubjects rather than :BookSubjects for the sake of simplicity }
\begin{DL}
$A \equiv Concept \sqcap B$ \\
$B \equiv \forall inScheme . C$ \\
$C \equiv ConceptScheme \sqcap ( \{BookSubjects\} \sqcup \{BookTopics\} \sqcup \{BookCategories\} )$
\end{DL}

@@

\begin{ex}
# valid data
# ----------
:ArtficialIntelligence
    a swrc:Book ;
    dcterms:subject :ComputerScience .
:ComputerScience
    a skos:Concept ;
    dcam:memberOf :BookSubjects ;
    skos:inScheme :BookSubjects .
:BookSubjects
    a skos:ConceptScheme .
\end{ex}

\begin{ex}
# invalid data
# ------------
:ArtficialIntelligence
    a swrc:Book ;
    dcterms:subject :ComputerScience .
:ComputerScience
    a skos:Concept ;
    dcam:memberOf :BooksAboutBirds ;
    skos:inScheme :BooksAboutBirds ;
    dcam:memberOf :BookSubjects ;
    skos:inScheme :BookSubjects .
:BookSubjects
    a skos:ConceptScheme .
\end{ex}

\begin{itemize}
	\item The related subject (\ms{:ComputerScience}) is a member of a controlled vocabulary (\ms{:BooksAboutBirds}) 
which is not part of the list of allowed controlled vocabularies.
\end{itemize}

>>>>

\ms{Not Allowed Values} corresponds to the requirements
\ms{R-33-NEGATIVE-OBJECT-} \ms{CONSTRAINTS} and
\ms{R-200-NEGATIVE-LITERAL-CONSTRAINTS}.
A matching triple has any literal / object except those explicitly excluded.

@@

Siths do not have blue, green, or white laser swords.

\begin{ex}
# OWL2:
laserSwordColor rdfs:range negativeLaserSwordColors . 
NegativeLaserSwordColors
    a rdfs:Datatype .
    owl:complementOf laserSwordColors .
laserSwordColors
    a rdfs:Datatype .
    owl:oneOf ( "blue" "green" "white" ) .
\end{ex}

\begin{ex}
# ShEx:
Sith {
    ! laserSwordColor ('blue' 'green' 'white') }
\end{ex}

\begin{ex}
# Sith individuals:
DarthSidious
    laserSwordColor 'red' .
\end{ex}

@@

\ms{Membership in Controlled Vocabularies} corresponds to the requirements \\
\ms{R-32-MEMBERSHIP-OF-RDF-OBJECTS-IN-CONTROLLED-VOCABULARIES} and \\
\ms{R-39-MEMBERSHIP-OF-RDF-LITERALS-IN-CONTROLLED-VOCABULARIES}
Resources can only be members of listed controlled vocabularies.

@@

\begin{ex}
# DSP:
bookDescriptionTemplate 
    a dsp:DescriptionTemplate ;
    dsp:resourceClass swrc:Book ; 
    dsp:statementTemplate [
        a dsp:NonLiteralStatementTemplate ;
        dsp:property dcterms:subject ; 
        dsp:nonLiteralConstraint [ 
            a dsp:NonLiteralConstraint ;
            dsp:valueClass skos:Concept ; 
            dsp:vocabularyEncodingScheme BookSubjects, BookTopics, BookCategories ] ] .
\end{ex}

\ms{skos:Concept} resources can only be members of the listed controlled vocabularies.

%\er{This sort of DL-syntax (with semicolon) must be rechecked. Note that in DL, it does not have to correspond directly to names of things in machine-readable syntax. It is fine to say BookSubjects rather than :BookSubjects for the sake of simplicity }
\begin{DL}
$A \equiv Concept \sqcap B$ \\
$B \equiv \forall inScheme . C$ \\
$C \equiv ConceptScheme \sqcap ( \{BookSubjects\} \sqcup \{BookTopics\} \sqcup \{BookCategories\} )$
\end{DL}

@@

\begin{ex}
# valid data:
ArtficialIntelligence
    a swrc:Book ;
    dcterms:subject ComputerScience .
ComputerScience
    a skos:Concept ;
    dcam:memberOf BookSubjects ;
    skos:inScheme BookSubjects .
BookSubjects
    a skos:ConceptScheme .
\end{ex}

\begin{ex}
# invalid data:
ArtficialIntelligence
    a swrc:Book ;
    dcterms:subject ComputerScience .
ComputerScience
    a skos:Concept ;
    dcam:memberOf BooksAboutBirds ;
    skos:inScheme BooksAboutBirds ;
    dcam:memberOf BookSubjects ;
    skos:inScheme BookSubjects .
BookSubjects
    a skos:ConceptScheme .
\end{ex}

The related subject (\ms{ComputerScience}) is a member of a controlled vocabulary (\ms{BooksAboutBirds}) 
which is not part of the list of allowed controlled vocabularies.
